---
title: "Parasite species distribution modeling"
author: "Tad Dallas, Andrew Park, and John M. Drake"
output:
  html_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    highlight: tango
    theme: journal
  pdf_document: default
---


```{r, echo=FALSE, message=FALSE, comment=FALSE}
#fish=read.csv('/media/drakelab/Lexar/HostParasiteRCN/Data/Fish/parasiteTraits.csv')
#load('/media/drakelab/Lexar/8910Project/Analysis/fishParty(reduced).RData')
#load('/media/drakelab/Lexar/8910Project/Analysis/fishParty.RData')
load('/media/drakelab/Lexar/8910Project/Analysis/fishParty(25min).RData')

library(rfishbase); library(gbm); library(ROCR); library(e1071); library(randomForest)
```


## Introduction




### Knowledge gap
What constrains the range of hosts that a parasite can infect? Is there a simple range of host functional traits that can determine the likelihood that a parasite infects a given host species?

Here, I develop predictive models based on host functional traits, and geographic location, to determine the host breadth for a large set of parasites of freshwater fish. 




### thorns in my side:

* Validation, validation, validation.
* Absence data aren't true absences. Should I even train on these data if the model treats them as true absences?
* 





## Methods

### Data and processing
 what the data are (fish-parasite data obtained from FishBase, and Strona's FishPest database)
 data selection, row removal, and imputation
 
 
### Models used
 discuss null predictions scenario, and then go into other algorithms used (brt, svm, lr, rf)




```{r, echo=FALSE, eval=FALSE}
#find all unique host species
fresh=fish[which(fish[,1] %in% unique(fish[,1])[c(1,4,6,7)]),]
marine=fish[which(fish[,1] %in% unique(fish[,1])[c(2,3)]),]

freshfish=unique(fresh$Host.fish.species); marinefish=unique(marine$Host.fish.species)
freshIndices=findSpecies(freshfish); marineIndices=findSpecies(marinefish)

# Get trait information
freshtraits=getQuantTraits(fish.data[freshIndices])
marinetraits=getQuantTraits(fish.data[marineIndices])

freshsize=cbind(getSize(fish.data[freshIndices], value='age'), getSize(fish.data[freshIndices], 'weight'), getSize(fish.data[freshIndices], 'length')); colnames(freshsize)=c('age', 'weight', 'length')

marinesize=cbind(getSize(fish.data[marineIndices], value='age'), getSize(fish.data[marineIndices], 'weight'), getSize(fish.data[marineIndices], 'length')); colnames(marinesize)=c('age', 'weight', 'length')

freshdepth=getDepth(fish.data[freshIndices])
marinedepth=getDepth(fish.data[marineIndices])

#format some trait matrices, and then remove any variables missing more than 90% observations
FT = cbind(freshtraits, freshsize, freshdepth)
FT = FT[,-which(colSums(is.na(FT)) > 500)]
MT = cbind(marinetraits, marinesize, marinedepth)
MT = MT[,-which(colSums(is.na(MT)) > 700)]

rownames(FT)=sub('_', ' ', rownames(FT))
rownames(MT)=sub('_', ' ', rownames(MT))
```





```{r, echo=FALSE, eval=FALSE}
#Data processing and support functions
edge2matrix=function(edgelist){
 return(table(edgelist[,1], edgelist[,2]))
}

# takes a matrix and returns an edgelist, where multiple reports are weighted by the number of entries in the resulting edgelist
matrix2edge=function(matrix, quant=TRUE){
   if(quant==FALSE){return(which(matrix>0, arr.ind=TRUE))
   }else{
    max=max(matrix)
    ret=list()
    for(i in 1:max){
      inds=which(matrix == i, arr.ind=TRUE)
      inds2=do.call("rbind", replicate(i, inds, simplify = FALSE))
      those=matrix(c(rownames(matrix)[inds2[,1]], colnames(matrix)[inds2[,2]]), ncol=2)
      ret[[i]]=those
    }
    ret2=matrix(unlist(ret), ncol=2, byrow=FALSE)
    colnames(ret2)=c('Host', 'Parasite')
    return(ret2)
   }
}


freshParasite = paste(gsub(' ', '', fresh[,3]), gsub(' ','', fresh[,4]), sep=' ')
freshParasite= gsub('Â', '', freshParasite)
freshHosts=sub('_', ' ',rownames(FT))

marineParasite = paste(gsub(' ', '', marine[,3]), gsub(' ','', marine[,4]), sep=' ')
marineParasite = gsub('Â', '', marineParasite)
marineHosts=sub('_', ' ', rownames(MT))

freshPotentialHosts=gsub('Â', '', fresh[,'Host.fish.species'])
freshPotentialHosts=gsub("^\\s+|\\s+$", "", freshPotentialHosts) 

marinePotentialHosts=gsub('Â', '', marine[,'Host.fish.species'])
marinePotentialHosts=gsub("^\\s+|\\s+$", "", marinePotentialHosts)

freshMatrix=edge2matrix(data.frame(parasite=freshParasite, host=freshPotentialHosts))
marineMatrix=edge2matrix(data.frame(parasite=marineParasite, host=marinePotentialHosts))

#set threshold of presence points to consider (n=5)
freshMatrix2=freshMatrix[which(rowSums(freshMatrix>0) > 25),]
freshMatrix3=freshMatrix2[, which(colnames(freshMatrix2) %in% freshHosts)]

marineMatrix2=marineMatrix[which(rowSums(marineMatrix>0) > 25),]
marineMatrix3=marineMatrix2[, which(colnames(marineMatrix2) %in% marineHosts)]



getPointsObject=function(parRow, hostTraits, binary=TRUE){
  hostTraits=hostTraits[order(rownames(hostTraits)),]
  inds=(rownames(hostTraits) %in% names(which(parRow>0)))
  occVec = rep(0, nrow(hostTraits))
  occVec[inds] = parRow[which(parRow>0)]
  if(binary==TRUE){occVec=0+(occVec>0)}
  return(data.frame(presence = occVec, hostTraits)) 
}


```







```{r, message=FALSE, eval=FALSE, echo=TRUE}
# list of imputed data matrices
imputedData=list()

# storage for model outputs
baseline.auc=vector(); allone.auc=vector(); allzero.auc=vector()
brtModel=list(); brt.best.iter=list(); brt.preds=list(); brt.perf=list(); brt.perfAUC=vector()
svmModel=list(); svm.preds=list(); svm.perf=list(); svm.perfAUC=vector()
lrModel=list(); lr.preds=list(); lr.perf=list(); lr.perfAUC=vector()
rfModel=list(); rf.preds=list(); rf.perf=list(); rf.perfAUC=vector()


for(i in 1:nrow(freshMatrix3)){
#data imputation
  dat = getPointsObject(freshMatrix3[i,], FT)
  dat = rfImpute(dat[,-1], dat[,1])
  imputedData[[i]] = dat
  
  pres=which(dat[,1]==1)
  inds=sample(1:nrow(dat), 0.7*nrow(dat))
  if(sum(pres %in% inds) < 4){inds[1:4]=pres[1:4]}
  
  train = dat[inds,]
  test = dat[-inds,]
  
  #baseline expectations and null models
  baseline.auc[i] = performance(prediction(test[sample(1:nrow(test), nrow(test)),1], test[,1]), 'auc')@y.values
   
 ##trained models
  #boosted regression trees
  weights=1/(sum(train[,1])/nrow(train))
  brtModel[[i]] = gbm(train[,1] ~ ., data=train[,-1], n.trees=60000, interaction.depth=4, distribution='bernoulli', weights=1+(train[,1] * weights))
  brt.best.iter[[i]] = gbm.perf(brtModel[[i]], method="OOB")
  brt.preds[[i]] = prediction(predict(brtModel[[i]], newdata=test[,-1], n.trees=brt.best.iter[[i]]), test[,1])
    brt.perf[[i]] = performance(brt.preds[[i]],"tpr","fpr")
  brt.perfAUC[i]=unlist(performance(brt.preds[[i]], 'auc')@y.values)

  #support vector machines
  svmModel[[i]] = svm(train[,1] ~ ., data=train[,-1], cross=5, probability=TRUE)
  svm.preds[[i]] = prediction(predict(svmModel[[i]], test[,-1]), test[,1])
  svm.perf[[i]] = performance(svm.preds[[i]],"tpr","fpr")
  svm.perfAUC[i] = unlist(performance(svm.preds[[i]], 'auc')@y.values)

  #logistic regression
  lrModel[[i]] = glm(train[,1] ~ ., data=train[,-1], family=binomial)
  lr.preds[[i]] = prediction(predict(lrModel[[i]], test[,-1]), test[,1])
  lr.perf[[i]] = performance(lr.preds[[i]],"tpr","fpr")
  lr.perfAUC[i] = unlist(performance(lr.preds[[i]], 'auc')@y.values)

  #random forest
  rfModel[[i]] = randomForest(train[,1] ~ ., data=train[,-1])
  rf.preds[[i]] = prediction(predict(rfModel[[i]], test[,-1]), test[,1])
  rf.perf[[i]] = performance(rf.preds[[i]],"tpr","fpr")
  rf.perfAUC[i] = unlist(performance(rf.preds[[i]], 'auc')@y.values)
  
  print(i)
}


```





```{r eval=TRUE, echo=TRUE, fig.cap='Mean predictive accuracy, measured as area under the ROC curve, for our null expectation (Base), boosted regression trees (BRT), support vector machines (SVM), logistic regression (LR), and random forest models (RF).'}
baseline.auc=unlist(baseline.auc)
#allzero.auc=unlist(allzero.auc)
#allone.auc=unlist(allone.auc)

aucs=cbind(baseline.auc, brt.perfAUC, svm.perfAUC, lr.perfAUC, rf.perfAUC)
colnames(aucs)=c('base', 'brt', 'svm', 'logr', 'rf')

meanAUC=colMeans(aucs)
meanSE = apply(aucs,2,sd) / sqrt(nrow(aucs))
meanSD = apply(aucs,2,sd)
  
plot(1:5, meanAUC, ylim=c(0.45, 1), xlim=c(0.5,5.5), las=1, pch=16, tck=0.01, xaxt='n', xlab='Predictive model used', ylab='Area Under ROC curve')
abline(h=0.5, col=grey(0.5), lty=2, lwd=2)
segments(x0=1:5, y0=meanAUC+meanSD, y1=meanAUC-meanSD, col=grey(0.8),lwd=2)
points(1:5, meanAUC, pch=16, cex=1.5)
axis(1, at=1:5, labels=c('Base', 'BRT', 'SVM', 'LR', 'RF'), tck=0.01, srt=30)

```





```{r, message=FALSE, eval=FALSE, echo=TRUE}
# list of imputed data matrices
m.imputedData=list()

# storage for model outputs
m.baseline.auc=vector(); m.allone.auc=vector(); m.allzero.auc=vector()
m.brtModel=list(); m.brt.best.iter=list(); m.brt.preds=list(); m.brt.perf=list(); m.brt.perfAUC=vector()
m.svmModel=list(); m.svm.preds=list(); m.svm.perf=list(); m.svm.perfAUC=vector()
m.lrModel=list(); m.lr.preds=list(); m.lr.perf=list(); m.lr.perfAUC=vector()
m.rfModel=list(); m.rf.preds=list(); m.rf.perf=list(); m.rf.perfAUC=vector()



for(i in 1:nrow(marineMatrix3)){
#data imputation
  dat = getPointsObject(marineMatrix3[i,], MT)
  dat = rfImpute(dat[,-1], dat[,1])
  m.imputedData[[i]] = dat
    
  pres=which(dat[,1]==1)
  inds=sample(1:nrow(dat), 0.7*nrow(dat))
  if(sum(pres %in% inds) < 4){inds[1:4]=pres[1:4]}
  train = dat[inds,]
  test = dat[-inds,]
  
  #baseline expectations and null models
  m.baseline.auc[i] = performance(prediction(test[sample(1:nrow(test), nrow(test)),1], test[,1]), 'auc')@y.values
   
 ##trained models
  #boosted regression trees
  weights=1 / (sum(train[,1])/nrow(train))
  m.brtModel[[i]] = gbm(train[,1] ~ ., data=train[,-1], n.trees=60000, interaction.depth=4, distribution='bernoulli', weights=1+(train[,1] * weights))
  m.brt.best.iter[[i]] = gbm.perf(m.brtModel[[i]], method="OOB")
  m.brt.preds[[i]] = prediction(predict(m.brtModel[[i]],newdata=test[,-1],n.trees=m.brt.best.iter[[i]]), test[,1])
  m.brt.perf[[i]] = performance(m.brt.preds[[i]],"tpr","fpr")
  m.brt.perfAUC[i]=unlist(performance(m.brt.preds[[i]], 'auc')@y.values)

  #support vector machines
  m.svmModel[[i]] = svm(train[,1] ~ ., data=train[,-1])
  m.svm.preds[[i]] = prediction(predict(m.svmModel[[i]], test[,-1]), test[,1])
  m.svm.perf[[i]] = performance(m.svm.preds[[i]],"tpr","fpr")
  m.svm.perfAUC[i] = unlist(performance(m.svm.preds[[i]], 'auc')@y.values)

  #logistic regression
  m.lrModel[[i]] = glm(train[,1] ~ ., data=train[,-1], family=binomial)
  m.lr.preds[[i]] = prediction(predict(m.lrModel[[i]], test[,-1]), test[,1])
  m.lr.perf[[i]] = performance(m.lr.preds[[i]],"tpr","fpr")
  m.lr.perfAUC[i] = unlist(performance(m.lr.preds[[i]], 'auc')@y.values)

  #random forest
  m.rfModel[[i]] = randomForest(train[,1] ~ ., data=train[,-1])
  m.rf.preds[[i]] = prediction(predict(m.rfModel[[i]], test[,-1]), test[,1])
  m.rf.perf[[i]] = performance(m.rf.preds[[i]],"tpr","fpr")
  m.rf.perfAUC[i] = unlist(performance(m.rf.preds[[i]], 'auc')@y.values)
  
  print(i)
}



```



```{r eval=TRUE, echo=TRUE}
m.baseline.auc=unlist(m.baseline.auc)
#m.allzero.auc=unlist(m.allzero.auc)
#m.allone.auc=unlist(m.allone.auc)

m.aucs=cbind(m.baseline.auc, m.brt.perfAUC, m.svm.perfAUC, m.lr.perfAUC, m.rf.perfAUC)
colnames(m.aucs)=c('base', 'brt', 'svm', 'logr', 'rf')

m.meanAUC=colMeans(m.aucs)
m.meanSE = apply(m.aucs,2,sd) / sqrt(nrow(m.aucs))
m.meanSD = apply(m.aucs,2,sd)
  
plot(1:5, m.meanAUC, ylim=c(0.45, 1), xlim=c(0.5,5.5), las=1, pch=16, tck=0.01, xaxt='n', xlab='Predictive model used', ylab='Area Under ROC curve')
abline(h=0.5, col=grey(0.5), lty=2, lwd=2)
segments(x0=1:5, y0=m.meanAUC+m.meanSD, y1=m.meanAUC-m.meanSD, col=grey(0.8),lwd=2)
points(1:5, m.meanAUC, pch=16, cex=1.5)
axis(1, at=1:5, labels=c('Base', 'BRT', 'SVM', 'LR', 'RF'), tck=0.01, srt=30)

```





## Results

## Discussion
